---
title: "Lab1_inR_NEllermeier"
output: html_notebook
---

#Biostat 705 Lab 1
#Kenya Datset
#Nathan Ellermeier
#Nine Eleven


#libraries
```{r}
library(haven)
library(skimr)
library(tidyverse)
library(ggplot2)
```


#pull in dataset
```{r}

kenya_dataset <- read_dta("data/lab_1_kenya.dta")
```


#explore dataset
```{r}
head(kenya_dataset)
```


```{r}
skim(kenya_dataset)
```

#creating new variable bord5 -- should we be creating these values as strings or numeric?
#if dichtomous, how to make dichotomous? in R, is that just a string of 0/1?
```{r}
kenya <- kenya_dataset %>%
  mutate(bord5 = NA)
    
kenya <- kenya %>%
  mutate(bord5 = case_when(
    bord <5 ~ "0",
    bord >=5 ~ "1",
  TRUE~NA_character_
  ))
  
```

```{r}
skim(kenya$bord5)

table(kenya$bord5)

```
#investigate b4
```{r}
skim(kenya$b4)
```


#generate male variable
#in original dataset, b4 is coded as 2 = woman and 1 = man
#NEED TO LABEL VARIABLES
```{r}
kenya <- kenya %>%
  mutate(male = NA)
    
kenya <- kenya %>%
  mutate(male = case_when(
    b4 ==2 ~ "0",
    b4 ==1 ~ "1",
  TRUE~NA_character_
  ))
```


#generating male2 using if_else command
#just for fun w/ if_else() command
```{r}
kenya <- kenya %>%
  mutate(male2 = if_else(b4 ==2, "0", "1"))
  
```

#variable labels
#fucking finally
#but all this did is created a vector and replaced my values with strings female and male  ????
```{r}
kenya$male2 <- factor(kenya$male2,
                levels = c(0,1),
                labels = c("Female","Male"))
```


#generate new variable mweight
```{r}
kenya <- kenya %>%
  mutate(mweight = v437/10)

```

#generate new variable mheight
```{r}
kenya <- kenya %>%
  mutate(mheight = v438/1000)
```

#generate new variable mbmi
```{r}
kenya <- kenya %>%
  mutate(mbmi = mweight / mheight)
```

```{r}
skim(kenya$mheight)
```



#now time to consider how we might save the following histograms so that it's easy for Larry
#Mother's Weight
```{r}
hist(kenya$mweight, main = "Mother's Weight" , xlab = "Mother's Weight in Kg", ylab = "# of occurences")
```


#I really wanna do some ggplot
#Mother's Height

```{r}
ggplot(data = kenya, aes(kenya$mheight)) +
         geom_histogram()
  
```

#different format
```{r}
kenya %>%
  ggplot() +
  geom_histogram(mheight, aes(color = "orange"))
```



#and compare it to base R hist()
```{r}
hist(kenya$mheight)
```


#qplot w/ mheight

```{r}
qplot(kenya$mheight, geom="histogram")
```

#qplot w/ adjustments
#argument for bindwidth either doesn't show perceivable change or give you a giant blue box over whole visualization-- wtf
#difference b/w I() and !I() in fill = and col = -- without it, the qplot() function prints a legend saying that "col="red""
#xlim = adjust the limit of the x axis with vector running between c(a,b)
#binwidth function does not seem to change binwidth!!
```{r}
qplot(kenya$mheight, 
      geom="histogram",
      bindwidth=.001,
      main = "Histogram of Mother's Height",
      xlab = "Height in Meters",
      ylab = "# of occurences",
      fill =I("blue"),
      col=I("red"),
      alpha=I(.2),
      xlim = c(1.2, 2)
      )
```




#ggplot w/ mheight
#ADDING
#why is this showing up differently for # of occurences than the histogram above?
#breaks() are used for binwidth. can also express the seq() argument using c(a,b) instead of seq(a,b, by=c)

#result renders "removed 308 rows containing non-finite values(stat_bin)"
#and also "removed 191 rows containing missing values (geom_bar)"
#what is the distinction here-- does data need cleaning for non-finite values? how to do that--

#aes(fill=..count..) renders a hist with darker colors for few fills and ligther colors for more fills
```{r}
ggplot(data=kenya, aes(x=mheight)) +
  geom_histogram(breaks=seq(0, 3, by=.02),
                 col="red",
                 aes(fill=..count..),
                 alpha = 1) +
  scale_fill_gradient("# of occurences", low="green", high="red") +
  labs(title="Histogram of Mother's Height", x="Height in Meters", y="# of occurences") +
  xlim(c(1,2.1)) +
  ylim(c(0,2000))
                 
```












#create histogram
```{r}
hist(kenya$bord , main = "Birth Order", xlab = "Age", ylab = "Number")
```


```{r}

```


































